---
title: "Assignment 1"
author: "Group 15: Ivona Bîrlad, Kaiyi Wang, Matteo Rapa"
date: "9/14/2022"
output: pdf_document
fontsize: 11pt
highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercise 1.1 Birthweight


```{r load data}
#load data
raw_data = read.table(file = "birthweight.txt", header = TRUE)
#check if the data is correctly loaded
head(raw_data)
birthweight = raw_data[, 1]
```
### a) Normality Check and $\mu$
```{r Normality check,  fig.cap="Normality Visualisation", fig.height=4, fig.width=8}
# histogram
par(mfrow= c(1,2))
hist(birthweight, main = "Histogram of Birth Weights", xlab = "Birth weight",  col="gray88", border="gray60")
# QQ Plot
qqnorm(birthweight, main = "Q-Q Plot of Birth Weights")
qqline(birthweight, col = "tomato", lwd = 2)
```

**Comments**: From data visualization, the data is normally distributed

```{r shapiro test}
# shapiro test for the normality of the data
shapiro.test(birthweight)
```

**Comments**: Reject H0, the data is normally distributed

```{r mu point estimate}
# the sample mean is used for point estimate for the data
bw_m = mean(birthweight); bw_m
```
**Comments**:

### b) Confidence interval for $\mu$

```{r CI of mu}
bw_n = length(birthweight)
bw_s = sd(birthweight)
# alpha/2 = 0.05, so 1 - alpha = 0.9
bw_t = qt(0.95, df = bw_n-1)
c(bw_m - bw_t*bw_s/sqrt(bw_n), bw_m + bw_t*bw_s/sqrt(bw_n))

```
**Comments**: The 90% CI of $\mu$ is: [2829.20, 2997.38] 

### c) one sample t-test
```{r one sample t-test}
# H0
# H1
# One sample t-test
t.test(birthweight, mu = 2800, alt = "g")
```
**Comments**: p-value is smaller than 0.05, here we reject H0. 

### d) further discussion
**Comments**: the R-output of the test from b), indicates that we are 90% confident that the interval from 2829.20 to 2997.38 actually does contain the true value of the population mean $\mu$. This CI is double side. In c), a single side test is conducted for the claim that the mean birthweight is bigger than 2800. To clarify this claim, we compute the alt = 'great' in the t-test. ....


## Exercise 1.2 Kinderopvangtoeslag
### a) estimate for p
```{r}
# point estimate for p
childcare_p = 140 / 200; childcare_p
```
**Comments**: the point estimate for p is 0.7

### b) CI of p
```{r}
# calculate q
childcare_q = 1 - childcare_p
# sample size
childcare_n = 200
# calculate z alpha/2. 
# for 99% CI, we have alpha = 0.01
childcare_z = qnorm(1 - 0.01/2)
# calculate the 99% CI of childcare_p
c(childcare_p - childcare_z * sqrt(childcare_p * childcare_q / childcare_n),
  childcare_p + childcare_z * sqrt(childcare_p * childcare_q / childcare_n))
```
**Comments**: The 99% CI of p is: [0.617, 0.783] 


### b) proportion test
```{r options(digits = 5)  }
bi_0.9 = binom.test(140, 200, p = 0.75,conf.level = 0.9); bi_0.9[3]
bi_0.95 = binom.test(140, 200, p = 0.75,conf.level = 0.95); bi_0.95[3]
bi_0.99 = binom.test(140, 200, p = 0.75,conf.level = 0.99); bi_0.99[3]
bi_0.8 = binom.test(140, 200, p = 0.75,conf.level = 0.8); bi_0.8[3]
```

Table 1 XXXX
$\alpha$ | 0.8 | 0.9 | 0.95 | 0.90 
------ | -------------------- | -----------------------  | -----------------------| ----------------------- 
P-value | `r  0.103` | `r 0.103`  | `r 0.103` | `r 0.103`

**Comments**: p-value doesn't change by different CI. Explanation?



## Exercise 1.3 Weather
```{r}
#load data
raw_data_weather = read.table(file = "weather.txt", header = TRUE)
#check if the data is correctly loaded
head(raw_data_weather)
# assign variables
humidity = raw_data_weather$humidity
temperature = raw_data_weather$temperature
```


```{r}
summary(humidity)
summary(temperature)
```



```{r,, fig.height=4, fig.width=8}
par(mfrow= c(1,2))
boxplot(humidity, main = "humidity", xlab = "humidity", ylab = "Percentage", col="dodgerblue")
boxplot(temperature, main = "temperature", xlab = "temperature", ylab = "fahrenheit", col="tomato")
```




```{r, fig.cap="Normality Visualisation", fig.height=4, fig.width=8}
# histogram
par(mfrow= c(1,2))
hist(humidity, main = "Histogram of humidity", xlab = "humidity",  col="gray88", border="gray60")
# QQ Plot
qqnorm(humidity, main = "Q-Q Plot of humidity")
qqline(humidity, col = "tomato", lwd = 2)
```
```{r shapiro test2}
# shapiro test for the normality of the data
shapiro.test(humidity)
```



```{r, fig.cap="Normality Visualisation", fig.height=4, fig.width=8}
# histogram
par(mfrow= c(1,2))
hist(temperature, main = "Histogram of temperature", xlab = "temperature",  col="gray88", border="gray60")
# QQ Plot
qqnorm(temperature, main = "Q-Q Plot of temperature")
qqline(temperature, col = "tomato", lwd = 2)
```
```{r shapiro test3}
# shapiro test for the normality of the data
shapiro.test(temperature)
```
